---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# oopr

<!-- badges: start -->
[![R-CMD-check](https://github.com/bochotti/oopr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bochotti/oopr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This package was created to assist with writing classes in R code. 
`methods::setRefClass()` and `R6::R6Class()` already provide such functionality,
but this package implements the idea differently.

  1. Class implementations are defined within a single brace `{`.

  2. Members can have different behavior via the use of specifiers.

  3. Upon creating a class, a series of checks are performed to reduce the
     risk of run-time errors.
     
  4. Code-completion and breakpoints are supported in RStudio.

Classes are useful for bundling data and actions together, and to "group" 
common functions.

## Installation

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("bochotti/oopr")
```

## Getting started

See `vignette("oopr")`.

## Example

```{r, include, echo=FALSE, results='asis'}
library(oopr);
text <- paste(readLines("./man/oopr.Rd"), collapse = "\n");
m <- regexpr("(?<=\\\\examples)(?<c>\\{([^}{]|(?&c))*\\})", text, perl = TRUE);
text <- regmatches(text, m);
text <- gsub("\\\\%", "%", text);
text <- sub("^\\{", "```{r}", text);
text <- sub("\\}$", "```",    text);
cat(knitr::knit_child(text = text, quiet = TRUE), sep = "\n");
```
