---
title: "Static"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Static}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Static members belong to a class itself, not the instances of a class. Static
members across multiple instances share the same state.

```{r setup}
library(oopr)
```

## Usage

Members can be declared static with the `static:` specifier.

```{r}
oopr("StaticExample",,
{
public:
  static:field  <- 1L;
})
```

```{r}
obj <- StaticExample();

# can be used from the class
print(StaticExample);
StaticExample$field <- 2L;

# instances share the same state with the class
print(obj);
```

## Requirements

  1. Static methods may only refer to other static members via `this$`.
  2. Static methods cannot be specified `virtual:`.
  
## Use-cases

### Global members

It can be useful for multiple instances to share the same variable:

```{r}
oopr("GlobalStaticExample",,
{
public:
  static:tally <- 0L;
  method <- \( ) { this$tally <- this$tally + 1L; }
})
```

```{r}
obj1 <- GlobalStaticExample();
obj2 <- GlobalStaticExample();

obj1$method();
print(obj2$tally);

obj2$method();
print(obj1$tally);
```

### Grouping common functions

Grouping common functions together in a single class can keep a namespace clean.

```{r}
oopr("UtilsExample",,
{
public:
  static:add      <- \(x, y) { return(x + y); }
  static:multiply <- \(x, y) { return(x * y); }
  static:subtract <- \(x, y) { return(x - y); }
})
```

```{r}
UtilsExample$add(1, 1);
UtilsExample$multiply(1, 1);
UtilsExample$subtract(1, 1);
```
