---
title: "Inheritance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inheritance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)
```

Inheritance is a way to re-ruse code across classes and establish a 
hierarchical relationship between classes. 

A relationship handles at least two classes: a base class and a derived class. 
The derived class *inherits* features from the base class, such as receiving the members of the derived class. Typically, inheritance justified when establishing
a *is a* relationship, i.e., the derived class *is a* base class.

```{r setup}
library(oopr)
```

## Usage

The second argument of `oopr()` is used for inheritance. It expects another 
class via its symbol. Multiple inheritance is available by using a brace `{`.

```{r}
# create some classes
oopr("Base1",, { public:shout   <- \( ) print("HELLO!"); })
oopr("Base2",, { public:whisper <- \( ) print("world");  })
```

```{r}
# single inheritance
oopr("InheritanceExample", Base1,
{
public:
  speak <- \( ) { this$shout(); }
})
obj <- InheritanceExample();
obj$speak();
```

```{r}
# multiple inheritance
oopr("InheritanceExample", { Base1; Base2; },
{
public:
  speak <- \( ) 
  { 
    this$shout(); 
    this$whisper(); 
  }
})
obj <- InheritanceExample();
obj$speak();
```

## Access

The access of base classes can be amended using `public:`, `protected:`,
and `private:` specifiers, this influences the access of all the base
classes members. By default, base classes are inherited privately.

If `public:`, then the base S3 class is included in the derived instance S3
class list.

```{r}
# Inherits private by default
print(obj)

# Inherit as public
oopr("InheritanceExample", public:Base1,
{
public:
  speak <- \( ) { this$shout(); }
})
obj <- InheritanceExample();

# Base public members are visible
print(obj);

# it also carries the S3 class
print(class(obj));
```

The interaction between member access and class inheritance access is:

| Member \\ Class | public    | protected | private   |
|:----------------|:----------|:----------|:----------|
| public          | public    | protected | private   |
| protected       | protected | protected | private   |
| private         | no access | no access | no access |

: Inheritance Access Interactions

This has some implications:

  1.  Public inheritance allows members of a derived class to match the 
      base class.
  2.  Protected members can be accessed by a derived class, but private
      members cannot.
  3.  Inheriting a class privately will allow use of its members to the 
      derived class, but hide them from further derived classes.


## Redefining Base Members

A derived class may create its own implementation of the base class method. 
The base class implementation can still be used inside a derived class 
methods via `base$` instead of `this$`:

```{r}
# base class with a method
oopr("Base",,
{
public:
  method <- \( ) { print("Base"); }
})
```

```{r}
# derived class, which redefines the method
oopr("Derived", public:Base,
{
public:
  method <- \( ) 
  { 
    Base$method();
    print("Method"); 
  }
})
```

```{r}
# only one method
obj <- Derived();
print(obj);

# which calls the base class
obj$method();
```

### Final

The `final:` specifier is available to stop derived classes redefining methods:

```{r}
oopr("Base",,
{
public:
  final:method <- \( ) { print("Base"); } #<- note final:
})
```

```{r}
oopr("Derived", public:Base,
{
public:
  method <- \( ) { print("Method"); }
})
```

## Virtual Methods

The base class may allow derived classes to implement their own versions of
a method, but require them to have the same function signature[^virtsig], 
which is useful for looping over multiple instances of different classes that 
all share a base class. 

[^virtsig]: As R is not type safe, just the arguments apply here, not the 
return type.

```{r}
# base class with virtual method
oopr("Base",,
{
public:
  virtual:method <- \(x, y = 1L) { }
})
```

```{r}
# doesnt match arguments
oopr("Derived", public:Base, 
{
public:
  method <- \( ) { }
})
```

```{r}
# doesnt have a default argument
oopr("Derived", public:Base, 
{
public:
  method <- \(x = 1L, y) { }
})
```

Virtual methods are *overridden* should a derived class choose to, I consider
this a form of "reverse inheritance". 

Although popular in C++, virtual methods in `oopr` cannot be specified as 
`private`.

```{r}
# method that relies on a virtual method
oopr("Base",,
{
public:
  speak  <- \( ) { this$method(); }
protected:
  virtual:method <- \( ) { print("Base"); }
})
```

```{r}
# now overridden
oopr("Derived", public:Base,
{
protected:
  method <- \( ) { print("Derived"); }
})
```

```{r}
obj <- Derived();
obj$speak()
```
