---
title: "Roxygen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Roxygen}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package exports `oopr_roclet()`, which is designed to help document
classes. To use additional roclets in a package workflow, refer to
`vignette("extending", "roxygen2")`.

## Layout

The roclet generates documentation across three main components: 

Constructor Method 
  : Appears in the **Usage** and **Arguments** sections (if applicable).

Value Section
  : Features a described list of each class, including its member count and
    description.
    
Class Sections 
  : Each class receives a dedicated section broken down 
    into the following hierarchy
    
  : Overview
    : Both **Description** and **Details** subsections.
    
  : Members
    : A described list of all **Fields** and **Methods**, including their 
      specifiers and descriptions.
      
  : Method Details
    : Each individual method breaks down into five subsections: 
      Description, Usage, Arguments, Details, and Returns.
    
See the `OoprVec()` documentation page for a live example of this layout.

## Usage 

### Classes

To document class-level `@description` and `@details`, use these tags directly
above the class definition. As each class has its own section in a page, it 
is recommended to set page-level documentation via `NULL`.

```{r, eval=FALSE}
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
#' @name example
#' @title example
#' @description
#' Page-level description
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
NULL

## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
#' @rdname example
#' @description
#' Class-level description
#'
#' @details
#' Class-level details
## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
oopr("example",, { })
```

### Members

All public members must be documented, if not a warning is thrown. Documenting
protected members is optional, and any documentation for private members is
ignored.

Members will have their specifiers automatically added to the documentation, so
users easily know if it is static, virtual, read-only, etc.

### Fields

Fields (and properties) can be documented by using the `@field` tag directly
above its definition:

```{r, eval=FALSE}
oopr("example",,
{
  ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
  #' @field field `integer(1L)` \cr
  #'              A documented field
  ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
  field <- 1L;
})
```

### Methods

Methods can be documented using the `@description`, `@usage`, `@param`, 
`@details`, and `@return` tags directly above the method definition.

```{r, eval=FALSE}
oopr("example",,
{
  ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
  #' @description
  #' A documented method
  #'
  #' @param x `integer(1)` \cr
  #'          A documented argument
  #'
  #' @details
  #' Some details on this method
  #'
  #' @returns
  #' The return value of this method
  ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ##
  method <- \(x)
  {
    return(x + this$field);
  }
})
```

A warning will be thrown if `@description`, `@returns` or `@param` (for an 
argument, if any) are not provided.

Any tags used for the constructor method are automatically moved to the 
page-level sections. A constructor method does not have its own section like
normal methods.

## Re-use

To avoid repeating the same documentation across members, they can be re-used
implicitly and explicitly.

### Implicit 

Missing tags trigger implicit documentation re-use based on two conditions:

Inherited Members
  : Automatically re-use documentation from the base class member.

Fields
  : Automatically re-use documentation from a constructor argument, if they 
    share the same name

### Explicit

Use the `@inherit` tag to pull the documentation of another class member 
(within from the same class or another). The expected syntax 
is `@inherit class$member`.

All inherited tags are copied over for methods, except those that are already
in use. Allowing a choice over which parts of the inherited member to use.

## .Rbuildignore

Generating documentation for internal classes can be helpful when using this
package, given the auto-complete functionality with RStudio with the
`devtools::load_all()` workflow. However, this generates a lot of help pages 
and aliases in the final package build which end-users do not need to see.

Any topic with an `oopr` class can use the `@keywords .Rbuildignore` tag to 
automatically add the `.Rd` file to `.Rbuildignore`. 

*Note* that the actual `.Rbuildignore` needs to be amended alongside removing 
the tag if one decides to stop build ignoring a `.Rd` file.

