---
title: "Class Members"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Class Members}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  error    = TRUE
)
```

Fields of a class can be defined as another class. When this happens the member 
becomes a class instance.

```{r setup}
library(oopr)
```

## Usage

Simply assign a class object to a member.

```{r}
oopr("MemberClass",, { public:x <- 1L; })
oopr("ClassMemExample",,
{
public:
  mem <- MemberClass;
})
```

```{r}
obj <- ClassMemExample();
print(obj);
```

## Containers

Members can become containers of classes using syntactic sugar. Suffixing `[]` 
uses `OoprVec()` and `[[]]` uses `OoprMap()`.

```{r}
oopr("ClassMemExample",,
{
ClassMemExample <- \( )
{
  for(i in 1:3)
  {
    this$mem$emplace();
    this$mem[i]$x <- i;
  }
}
public:
  mem <- MemberClass[]; # <- use of []
})
```

```{r}
obj <- ClassMemExample();
print(obj);
```

## Construction

If the class member has a constructor method, then the member will need to:

  1.  Be initialized within the constructor method, or
  2.  Have its arguments provided where the member is being defined.

If the member is specified as `static:`, then 2. is the only option.

```{r}
oopr("MemberClass",, 
{ 
MemberClass <- \(x) { this$x <- x; }
public:x    <- 1L; 
})

# not allowed, needs constructing
oopr("ClassMemExample",,
{
public:
  mem <- MemberClass;
})
```

```{r}
# both approaches
oopr("ClassMemExample",,
{
ClassMemExample <- \( )
{
  this$mem1(1L);           # <- args given in constructor
}
public:
  mem1 <- MemberClass;
  mem2 <- MemberClass(2L); # <- args given in-line
})
```

```{r}
obj <- ClassMemExample();
print(obj);
```
