---
title: "Inheritance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inheritance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Inheritance is a way to re-ruse code and establish a hierarchical relationship
between classes.

A relationship handles at least two classes: a base class and a derived class.
The derived class *inherits* features from the base class, such as receiving
the members of the derived class. Typically, inheritance is justified when
establishing an *is a* relationship, i.e., the derived class *is a* base class.

```{r setup}
library(oopr)
```

## Usage

The second argument of `oopr()` is used for inheritance. It expects another
class via its symbol. Multiple inheritance is available by using a brace `{`.

```{r}
# create some classes
oopr("Base1",, { public:shout   <- \( ) print("HELLO!"); })
oopr("Base2",, { public:whisper <- \( ) print("world");  })
```

```{r}
# single inheritance
oopr("InheritanceExample", Base1,
{
public:
  speak <- \( ) { this$shout(); }
})
```

```{r}
obj <- InheritanceExample();
obj$speak();
```

```{r}
# multiple inheritance
oopr("InheritanceExample", { Base1; Base2; },
{
public:
  speak <- \( )
  {
    this$shout();
    this$whisper();
  }
})
```

```{r}
obj <- InheritanceExample();
obj$speak();
```

## Access

The access of base classes can be amended using `public:`, `protected:`,
and `private:` specifiers, this influences the access of all the base
classes members. By default, base classes are inherited privately.

If `public:`, then the base S3 class is included in the derived instance S3
class list.

```{r}
# Inherits privately by default
print(obj)
```

```{r}
# Inherit as public
oopr("InheritanceExample", public:Base1,
{
public:
  speak <- \( ) { this$shout(); }
})
```

```{r}
obj <- InheritanceExample();

# Base public members are visible
print(obj);

# it also carries the S3 class
print(class(obj));
```

The interaction between member access and class inheritance access is:

| Member \\ Class | public    | protected | private   |
|:----------------|:----------|:----------|:----------|
| public          | public    | protected | private   |
| protected       | protected | protected | private   |
| private         | no access | no access | no access |

: Inheritance Access Interactions

This has some implications:

  1.  Public inheritance allows members of a derived class to copy the
      base class.
  2.  Protected members can be accessed by a derived class, but private
      members cannot.
  3.  Inheriting a class privately will allow use of its members to the
      derived class, but hide them from further derived classes.

Note that constructor and destructor methods are never inherited.

## Redefining Base Members

A derived class may create its own implementation of the base class method.
The base class implementation can still be used inside a derived class
methods via `base$` instead of `this$`:

```{r}
# base class with a method
oopr("Base",,
{
public:
  method <- \( ) { print("Base"); }
})
```

```{r}
# derived class, which redefines the method
oopr("Derived", public:Base,
{
public:
  method <- \( )
  {
    Base$method();
    print("Method");
  }
})
```

```{r}
# only one method
obj <- Derived();
print(obj);

# which calls the base class
obj$method();
```

### Final

The `final:` specifier is available to stop derived classes redefining methods:

```{r}
oopr("Base",,
{
public:
  final:method <- \( ) { print("Base"); } #<- note final:
})
```

```{r, error = TRUE, purl = FALSE}
# not allowed
oopr("Derived", public:Base,
{
public:
  method <- \( ) { print("Method"); }
})
```

## Virtual Methods

The base class may allow derived classes to implement their own versions of
a method, but require them to have the same function signature[^virtsig],
which is useful for looping over multiple instances of different classes that
all share a base class.

[^virtsig]: As R is not type safe, just the argument names apply, not the
return type.

```{r}
# base class with virtual method
oopr("Base",,
{
public:
  virtual:method <- \(x, y = 1L) { }
})
```

```{r, error = TRUE, purl = FALSE}
# does not match arguments
oopr("Derived", public:Base,
{
public:
  method <- \( ) { }
})
```

```{r, error = TRUE, purl = FALSE}
# does not have a default argument
oopr("Derived", public:Base,
{
public:
  method <- \(x = 1L, y) { }
})
```

Virtual methods inside the base class instance are *overridden* if the derived
class creates its own implementation of the method.

```{r}
# method that relies on a virtual method
oopr("Base",,
{
public:
  speak  <- \( ) { this$method(); }
protected:
  virtual:method <- \( ) { print("Base"); }
})
```

```{r}
# overridden for the derived class
oopr("Derived", public:Base,
{
protected:
  method <- \( ) { print("Derived"); }
})
```

```{r}
obj <- Derived();
obj$speak()
```

Although popular in other languages, virtual methods with `oopr` cannot be
specified as `private:`.

### Abstract

Virtual methods can be marked as abstract. An abstract method requires derived
classes to create their own implementation. Any class with an abstract method
cannot have an instance constructed.

To mark a virtual method as abstract, append `= 0` after its body.
```{r}
# mark an abstract method
oopr("AbstractClass",,
{
public:
  doWork <- \(x, y, z)
  {
    o <- this$doStep1(x, y);
    o <- this$doStep2(o, z);
    return(o);
  }
protected:
  virtual:doStep1 <- \(x, y) { } = 0 #<- = 0 here
  virtual:doStep2 <- \(o, z) { } = 0 #<- = 0 here
})
```

```{r, error = TRUE, purl = FALSE}
# Cannot create an instance
AbstractClass()

# Cannot be a class member
oopr("AbstractClassMem",,
{
public:
  abs <- AbstractClass
})

# Derived class cannot be final
oopr("AbstractClassInhr", public:AbstractClass,
{
public:
  final:AbstractClassInhr <- \( ) { }
})
```

```{r}
# correct
oopr("DerivedClass", public:AbstractClass,
{
protected:
  doStep1 <- \(x, y) { return(x + y); }
  doStep2 <- \(o, z) { return(o * z);   }
})

obj <- DerivedClass();
obj$doWork(1, 2, 3);
```

## Constructing

When the base class has a constructor method, then it must be explicitly
initialized in the derived class[^constr].

[^constr]: Unless the base class' constructor method has no arguments,
or all arguments have a default value.

```{r}
# base class with constructor method
oopr("Base",,
{
public:
  Base <- \(x) { this$x <- x; }
  x    <- integer(1L);
})
```

```{r, error = TRUE, purl = FALSE}
# unable to create the derived class
oopr("Derived", Base, {})
```

To initialize, simply call the base class constructor within the derived
customer method:

```{r}
# call Base(...)
oopr("Derived", public:Base,
{
public:
  Derived <- \(x) { Base(x); }
})
```

```{r}
# success
obj <- Derived(1L);
print(obj);
```

Note that base class members cannot be accessed until they are constructed:

```{r, error = TRUE, purl = FALSE}
# cannot access `x` before Base(...)
oopr("Derived", public:Base,
{
public:
  Derived <- \(x)
  {
    print(Base$x); # <- neither Base$
    print(this$x); # <- nor this$
    Base(x);
  }
})
```

### Protected Constructor

To ensure that a class can only be constructed as a base class, mark its
constructor method as `protected:`:

```{r}
# protected constructor
oopr("Base",,
{
protected:
  Base <- \( ) { }
})
```

```{r, error = TRUE, purl = FALSE}
# not allowed
Base();
```

```{r}
# allowed
oopr("Derived", Base, { })
Derived();
```


### Final Constructor

To prevent a class from being inherited by other classes, mark its constructor as `final:`:

```{r}
# final constructor
oopr("Base",,
{
public:
  final:Base <- \( ) { this$x <- 10L; }
  x <- integer(1L);
})
```

```{r, error = TRUE, purl = FALSE}
# not allowed
oopr("Derived", Base, { })
```

```{r}
# allowed
Base();
```

Note: private and protected constructors cannot be marked as final, as this
would not be intuitive.

