---
title: "Class Members"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Class Members}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

Fields of a class can be defined as another class. When this happens the member 
becomes a class instance.

```{r setup}
library(oopr)
```

## Usage

Simply assign a class object to a member.

```{r}
oopr("MemberClass",, { public:x <- 1L; })
oopr("ClassMemExample",,
{
public:
  mem <- MemberClass;
})
```

```{r}
obj <- ClassMemExample();
print(obj);
```

## Protection

Class members are automatically protected via an active binding. When assigning
into a class member, the incoming value is checked for the same class:

```{r, error = TRUE, purl = FALSE}
# allowed
obj$mem$x <- 2L;

# not allowed
obj$mem <- 1L;

# allowed
obj$mem <- MemberClass();

# not allowed
obj$mem <- ClassMemExample();
```

Using set properties (or the absence of one for read-only) with classes may 
not work as intended as R expressions are evaluated from right to left. An
expression such as `obj$mem$a <- 1L;` will evaluate `mem$a <- 1L;`
before the set property for `obj$mem` is called. 

A useful set property to ensure the class instance is never replaced could be:
```{r}
oopr("ClassMemPropertyExample",,
{
public:
  get:x <- \( ) { return(this$x_); }
  set:x <- \(rhs)
  {
    if(!identical(this$x_, rhs)) 
    { 
      stop(call. = FALSE, "`x` cannot be replaced"); 
    }
    return(this$x_);
  }
private:
  x_ <- MemberClass;
})
```

```{r, error = TRUE, purl = FALSE}
obj <- ClassMemPropertyExample();

# allowed
obj$x$x <- 2L;

# not allowed
obj$x <- MemberClass();
```

## Containers

Members can become containers of classes using syntactic sugar. Suffixing `[]` 
uses `OoprVec()` and `[[]]` uses `OoprMap()`.

```{r}
oopr("ClassMemExample",,
{
public:
  ClassMemExample <- \( )
  {
    for(i in 1:3)
    {
      this$mem$emplace();
      this$mem[i]$x <- i;
    }
  }
  mem <- MemberClass[]; # <- use of []
})
```

```{r}
obj <- ClassMemExample();
print(obj);
```

## Construction

If the class member has a constructor method, then the member will need to:

  1.  Be initialized within the constructor method, or
  2.  Have its arguments provided where the member is being defined.

If the member is specified as `static:`, then 2. is the only option.

```{r}
oopr("MemberClass",, 
{ 
public:
  MemberClass <- \(x) { this$x <- x; }
  x <- 1L; 
})
```

```{r, error = TRUE, purl = FALSE}
# not allowed, needs constructing
oopr("ClassMemExample",,
{
public:
  mem <- MemberClass;
})
```

```{r}
# both approaches
oopr("ClassMemExample",,
{
public:
  ClassMemExample <- \( )
  {
    this$mem1(1L);         # <- args given in constructor
  }
  mem1 <- MemberClass;
  mem2 <- MemberClass(2L); # <- args given in-line
})
```

```{r}
obj <- ClassMemExample();
print(obj);
```
